Revisiting the Relationship between Short-Chain Fatty Acids, the Microbiota, and Colorectal Tumors

### 2018-08-09 submission/manuscript.tex
final 1st draft read through

\textbf{Background.} Colorectal cancer (CRC) is increasing in prevalence
in individuals under 50 and because of this will be a continuing health
concern for the foreseeable future. The majority of the risk for
developing CRC is attributable to environmental factors. One of these
environmental factors is the microbiota, with certain bacterial
community members being associated with CRC and other taxa being
associated to individuals without tumors. Some of the bacterial species
in taxa associated to individuals without tumors can use fiber to
produce short-chain fatty acids (SCFAs) that inhibit tumor growth in
model systems. However, the data supporting the importance of SCFAs in
human CRC is less certain. Here, we test the hypothesis that SCFA
concentrations and taxa associated with their production are different
in individuals with colorectal tumors.

@@

\textbf{Conclusions.} Although our data does not support the hypothesis
that fecal SCFA concentrations in patients in the general CRC population
are different, there still may be specific types of colorectal tumors
where SCFAs may be beneficial for treatment of CRC. Alternatively, our
observations also support the hypothesis that there may be other
metabolites or mechanisms (e.g.~bacterial niche exclusion) that may be
more protective against tumorigenesis and have not been thoroughly
investigated in the context of human CRC.

@@

Our study fills some of the current gaps in the literature that relate
to the study of SCFAs and CRC in human populations. Specifically, by
using a seperate cohort it tests previous case/control findings on SCFA
concentrations in individuals with and without tumors. Additionally,
prior investigations grouped patients with adenoma and individuals
without tumors into a single group to compare against patients with
carcinomas. Despite doing this, the suggestion was made that a reduction
in SCFA concentration would also be observed between patients with
adenomas and indivdiuals without tumors. We also can test this
suggestion because we have a larger number of patients with adenomas
within our study and do not need to group patients with adenomas and
individuals without tumors into a single group. Additionally, we build
upon these observations by assessing the utility of using SCFAs and
Operational Taxonomic Units (OTUs) as a risk stratification tool of
colorectal tumors (adenoma or carcinoma). We also investigate whether
OTUs that are most important to these models are closely associated with
the classification of SCFA concentrations. Collectively, this study
provides important information on the replicability of previous findings
in humans by extensively studying how SCFAs are associated with
colorectal tumors.

To accomplish this task we directly measured the concentration of
acetate, butyrate, and propionate within fecal samples for two different
groups. The first group had a sample obtained at a single cross
sectional point in time while the second group had samples obtained
before (pre-) and after (post-)treatment for colorectal tumors. To
provide further support for our SCFA findings we also used PICRUSt (21)
and metagenomic sequencing to investigate differences in genes involved
with SCFA synthesis between individuals without tumors, patients with
adenomas, and patients with carcinomas. Next, we investigated whether
taxa associated with SCFA production were important to disease
classification models. First, using the cross-sectional data, we
analyzed the number of correlations between OTU relative abundance and
SCFA concentrations across individuals without tumors and patients with
adenomas or carcinomas. Second, we assessed the affect adding SCFA
concentrations to OTU data had on classification of patients with
adenomas or carcinomas using the Random Forest algorithm (22). Third, we
analyzed how well 16S rRNA gene sequencing predicts SCFA concentrations.
Finally, we compared whether SCFA concentrations replace important taxa
in disease models and whether the same OTUs are the most important
variables used to classify disease and SCFA concentration. Collectively,
this investigation provides additional information as to whether SCFAs
are decreased in patients with colorectal tumors and provides context as
to whether targeting taxa to increase SCFA concentrations is a viable
option to protect against colon tumorigenesis.

@@

\textbf{Gene abundance for enzymes involved in SCFA synthesis are the
same for individuals without tumors and patients with adenomas or
carcinomas.} In order to provide further confirmation and support of our
SCFA concentration results we investigated the genes encoding specific
enzymes involved with SCFA synthesis {[}Table S1{]}. Using this list of
specific genes {[}Table S1{]}, we looked for differences in gene
abundance between individuals without colon tumors and patients with
adenomas or carcinomas. Although we intended to analyze all the genes in
the list {[}Table S1{]}, not all of the KEGG genes were identified
during our analysis. We first analyzed imputed gene relative abundance,
calculated from our OTU data that was generated from 16S rRNA gene
sequencing. We found no difference in any of the imputed gene relative
abundances for enzymes involved with acetate, butyrate, or propionate
synthesis (P-value \textgreater{} 0.90) {[}Table S2{]}. Since butyrate
is one of the most studied SCFAs in the context of CRC, we visualized
the observed lack of difference between indviduals without tumors and
patients with adenomas or carcinomas in gene abundance for enzymes
involved with SCFA synthesis using the butyrate kinase gene {[}Figure
2A{]}. Additionally, using a paired Wilcoxon rank-sum test, there also
was no difference in imputed gene relative abundance between pre- and
post-treatment samples for any genes involved with SCFA synthesis
(P-value \textgreater{} 0.70) {[}Table S3{]}. Next, we took a subset of
these 490 fecal samples (n=85) and used metagenomic sequencing to
confirm these results. Like the imputed gene results, metagenomic
analysis found that there was no difference in any of the genes involved
in SCFA synthesis between individuals without colon tumors (n=29) and
patients with adenomas (n=28) or carcinomas (n=28) (P-value
\textgreater{} 0.70) {[}Table S4{]}. This similarity between indviduals
without tumors and patients with adenomas or carcinomas is highlighted
again by visualizing the gene abundance for the butyrate kinase gene
{[}Figure 2B{]}. These observations provide evidence that gene
prevalence for enzymes involved in SCFA synthesis does not change due to
colorectal tumors and provides further support for our original SCFA
concentration observations.

\textbf{The number of OTUs positively associated with SCFA concentration
were similar between individuals without tumors and patients with
adenomas or carcinomas.} Having found no difference between individuals
without tumors and patients with adenomas or carcinomas in SCFA
concentrations or genes encoding enzymes involved with SCFA synthesis,
we next investigated if specific OTUs correlated with SCFA
concentrations. The main goal of this analysis was to identify if there
were OTUs that were significantly associated with SCFA concentrations
and if this was different between individuals that did not have tumors,
had an adenoma, or had a carcinoma. To accomplish this we used Spearmans
rho, a non-parametric measure of association, and tested if there was a
correlation that was signficantly greater than zero. We found that taxa
from \emph{Clostridiales}, \emph{Lachnospiraceae}, and
\emph{Ruminococcaceae} dominated statistically significant OTU
correlations {[}Figure 3 \& Table S5{]}. There was a noticeably higher
number of significant negative correlations associated with patients
with adenomas for all SCFAs tested {[}Figure 3{]}. In particular, OTUs
from the \emph{Ruminococcaceae} family had the largest share of these
negative correlations within patients with adenomas {[}Figure 3{]}.
Patients with adenomas also had more positive correlations between OTUs
and SCFA concentrations, but their total number was more similar to
individuals without tumors or patients with carcinomas versus the
analogous comparison for the number of negative correlations {[}Figure
3{]}. Additionally, the number of positive correlations between OTUs and
SCFA concentrations was similar between individuals without tumors and
patients with a carcinoma {[}Figure 3{]}. Finally, when we used high/low
SCFA groups based on the overall median concentration for each SCFA
instead of SCFA concentrations a similar pattern was still observed
{[}Figure S1 \& Table S6{]}. Overall, these results suggest that the
resident taxa that may change the most due to colon tumors may not be
ones that are responsible for the production of acetate, butyrate, or
propionate.

\textbf{SCFA concentrations do not replace important
\emph{Clostridiales}, \emph{Lachnospiraceae}, and \emph{Ruminococcaceae}
OTUs in Random Forest models built to classify tumors.} Despite the lack
of difference in positive correlations between OTUs and SCFA
concentrations between individuals with and without tumors, OTUs
associated with SCFA concentrations could still be the most important
variables to Random Forest models built to classify patients with
adenomas or carcinomas. We tested this by using the Random Forest
algorithm to build models with OTU abundance data or OTU abundances and
SCFA concentrations to classify normal versus adenoma and normal versus
carcinoma fecal samples. With these models we compared whether OTUs with
taxonomic classification to \emph{Clostridiales},
\emph{Lachnospiraceae}, and \emph{Ruminococcaceae} remained when SCFA
concentrations were added to the model. Additionally, we also compared
whether any of the important OTUs that remained also had significant
correlations with SCFA concentrations. Both our adenoma and carcinoma
models classified patients with a similar degree of success, as measured
by the area under the curve (AUC) (P-value \textgreater{} 0.05)
{[}Figure 4A \& 4D{]}. After the addition of SCFA concentrations to the
adenoma or carcinoma models, many OTUs with taxonomic classification to
\emph{Clostridiales}, \emph{Lachnospiraceae}, and \emph{Ruminococcaceae}
remained as important variables to the model {[}Figure Figure 4B-C \&
4E-F{]}. After adding SCFA concentrations to the adenoma model, there
were only 2 OTUs significantly associated with SCFA concentration that
were part of the top 10 most important variables and both were postively
associated with acetate, butyrate, or propionate concentrations. When
SCFA concentrations were added to the carcinoma model, only 1 OTU that
was associated with SCFA concentration remained as part of the top 10
most important variables and it was negatively associated with acetate
and butyrate concentrations. In combination with the previous results on
OTU correlations, these observations provide additional evidence that
the resident taxa that are associated with protection against
tumorigeneis are not ones associated with acetate, butyrate, or
propionate production.

\textbf{The most important OTUs in Random Forest models built to
classify SCFA concentrations or tumors are not the same.} It is possible
that due to the way the Random Forest algorithm works, OTUs associated
with SCFA concentrations could be downweighted in importance within the
adenoma or carcinoma models when SCFA concentrations are included. To
test if this is the case we used OTU data and built Random Forest models
to classify SCFA concentrations. Overall, the correlation between the
predicted and actual SCFA concentrations were moderately associated with
each other {[}Figure 5A{]}. Additionally, because the training set
R\textsuperscript{2} was always higher than the test set
R\textsuperscript{2}, all SCFA concentration models tended to be over
fit, suggesting that rarer taxa were important for these classifications
{[}Figure 5A{]}. There also was a difference in accuracy based on
whether the fecal sample was from an individual without tumors or from
patients with adenomas or carcinomas {[}Figure 5B{]}. When comparing the
adenoma or carcinoma model to the SCFA concentration models there was
minimal overlap between these model's most important OTUs {[}Figure
4B-C, 4E-F and 5C-E{]}. The only OTU that was in the top 10 most
important variables and had overlap between the models was OTU00167
(\emph{Clostridiales}) {[}Figure 4B-C, 4E-F, 5C-E{]}. Similar
observations were made when using high/low SCFA groups based on the
median SCFA concentration {[}Figure S2{]}. Collectively, these
observations provide evidence that it is possible to identify specific
OTUs associated with higher SCFA concentrations and accordingly these
OTUs belong to taxa known to produce acetate, butyrate, and propionate.
Although it is possible to identify OTUs associated with SCFA
production, overall, our results do not support the hypothesis that SCFA
concentration or OTUs associated with their production are different
between individuals with no tumors and patients with adenomas or
carcinomas.

@@

The observations from this study do not support the hypothesis that SCFA
concentrations are different in individuals with tumors. Whether we
directly measured the SCFA concentration or investigated genes that
encoded enzymes used for their production, no difference could be
identified between individuals without tumors and patients with adenomas
or carcinomas {[}Figure 1 \& 2{]}. Although there were differences in
the number of significant correlations between SCFA concentration and
OTU relative abundance based on whether individuals did not have tumors,
had an adenoma, or had a carcinoma, SCFA concentrations did not provide
increased model accuracy for tumor classification {[}Figure 3-4 \&
S1{]}. In models with SCFA concentrations included, many OTUs that
classified to \emph{Clostridiales}, \emph{Lachnospiraceae}, and
\emph{Ruminococcaceae} remained as important variables for the model
{[}Figure 4{]}. Additionally, when models using OTU relative abundance
to classify SCFA concentrations were assessed, the OTUs that classified
to \emph{Clostridiales}, \emph{Lachnospiraceae}, and
\emph{Ruminococcaceae} were not the same as the OTUs that classified to
these taxa in the tumor models {[}Figure 4-5{]}. Collectively, our
observations suggest that the resident taxa from \emph{Clostridiales},
\emph{Lachnospiraceae}, and \emph{Ruminococcaceae} that are different
between individuals without tumors and patients with adenomas or
carcinomas, are not the same as those involved with SCFA production.

Although SCFAs have been shown to be anti-tumorigenic, most of these
studies have been performed in model systems (16, 17). Many of the
\emph{in vivo} studies use proxies such as fiber supplementation rather
than SCFAs directly (14). Although it is well known that breakdown
products from gut bacteria results in SCFA production, fiber effects on
tumorigenesis may be through other mechanisms in these \emph{in vivo}
model systems. Additionally, the observations in humans on the benefit
of SCFAs in preventing tumorigenesis have been mixed. In previous
case/control studies lower SCFA concentrations were observed in patients
with carcinomas versus those without carcinomas (19). Yet, this is in
contrast to multiple randomized-controlled trials that have found no
difference in tumor recurrence between patients who do and do not get
fiber supplementation (20, 23). In contrast to the \emph{in vivo} model
findings, the observations made in these randomized-controlled trials
would suggest that SCFAs do not prevent or slow tumorigenesis. One
reason for these results is that SCFA concentrations and responses to
fiber vary quite a bit between healthy individuals (24). This
information taken together with our observations would suggest that
either individuals who do not respond to fiber supplementation would
need to acquire these bacteria to achieve a benefit or that SCFAs
provide little to no benefit as an anti-tumorigenic compound in
colorectal cancer.

Another possible alternative explanation as to why no difference in SCFA
concentration between individuals without tumors and patients with
adenomas or carcinomas was observed, could be because only certain types
of colorectal cancers are affected by SCFAs. One limitation of current
research into the effect of SCFAs and the microbiota in CRC has been
that all tumors are treated as the same type. However, there are known
differences in the types of mutations that occur (25) and treating all
tumors as equal may actually hide any benefit that could be found in
certain subsets of individuals. Similar to the idea of using
immunotherapy as a targeted treatment option for specific tumors (26),
SCFAs may have beneficial effects for distinct types of colorectal
tumors. Future research will need to test if this is a valid hypothesis.
Regardless of this limitation, our results in combination to previous
randomized controlled trials on fiber supplementation (20, 23) suggests
that using SCFAs as a general treatment for colorectal cancer is
unlikely to provide a reduction in tumorigenesis.

@@

The potential protection against colorectal cancers may not be from
SCFAs, even though taxa associated with their production are higher in
individuals without tumors versus patients with carcinomas (10, 11, 27).
Our data is consistent with the literature despite finding no difference
in SCFA concentrations between individuals without tumors and patients
with adenomas or carcinomas. In particular, we observed that the
important OTUs in our tumor models were to taxa that are associated with
SCFA production but that these OTUs were not themselves associated with
acetate, butyrate, or propionate production. One of the clearest results
to highlight this within our study were the results showing that
different OTUs from the same taxonomic classification are in tumor and
SCFA Random Forest models {[}Figure 4 \& 5{]}. This leads to the
possibility that protection may be through other routes. First, there
could be a different pathway or other less extensively studied
metabolites that provide the necessary protection against tumorigenesis.
Alternatively, previous research has shown that non-resident oral
microbes, such as \emph{Fusobacterium}, \emph{Porphyromonas},
\emph{Parvimonas}, and \emph{Peptostreptococcus} are associated with
carcinomas (6, 12, 13) and can increase inflammation (28). When these
non-resident oral microbes increase inflammation they change the
existing environment to one in which is more hospitable for themselves
(28). It is possible that these resident microbes provide protection not
through a metabolite but instead through active exclusion of these
non-resident oral microbes which in turn prevents them from establishing
themselves within the community. Overall, even though we did not find
lower concentrations of SCFAs associated with colorectal tumors, we
think that our results open up many new and exciting avenues for
investigation into how metabolites and the microbiota can prevent
tumorigenesis.

>>>>

\textbf{Background.} Colorectal cancer (CRC) is increasing in prevalence
in individuals under 50 and because of this will be a continuing health
concern for the foreseeable future. The majority of the risk for
developing CRC is attributable to environmental factors. One of these
environmental factors is the microbiota, with certain bacterial
community members being more prevalent in CRC while other taxa are
higher in individuals without tumors. Some of the bacterial species in
taxa that are more abundant in individuals without tumors can use fiber
to produce short-chain fatty acids (SCFAs) that inhibit tumor growth in
model systems. However, the data supporting the importance of SCFAs in
human CRC is less certain. Here, we test the hypothesis that SCFA
concentrations and taxa associated with their production are different
in individuals with colorectal tumors.

@@

\textbf{Conclusions.} Although our data does not support the hypothesis
that fecal SCFA concentrations are different in the general CRC
population, there still may be specific types of colorectal tumors where
SCFAs may be beneficial for treatment of CRC. Alternatively, our
observations also support the hypothesis that there may be other
metabolites or mechanisms (e.g.~bacterial niche exclusion) that may be
more protective against tumorigenesis and have not been thoroughly
investigated in the context of human CRC.

@@

Our study fills some of the current gaps in the literature that relate
to the study of SCFAs and CRC in human populations. Specifically, by
using a seperate cohort it tests previous case/control findings on SCFA
concentrations in individuals with and without tumors. Additionally,
prior investigations grouped patients with adenoma and individuals
without tumors into a single group to compare against patients with
carcinomas. Despite doing this, the suggestion was made that a reduction
in SCFA concentration would also be observed between patients with
adenomas and indivdiuals without colon tumors. We also can test this
suggestion because we have a larger number of patients with adenomas
within our study and do not need to group patients with adenomas and
individuals without colon tumors into a single group. Additionally, we
build upon these observations by assessing the utility of using SCFAs
and Operational Taxonomic Units (OTUs) as a risk stratification tool of
colorectal tumors (adenoma or carcinoma). We also investigate whether
OTUs that are most important to these models are closely associated with
the classification of SCFA concentrations. Collectively, this study
provides important information on the replicability of previous findings
in humans by extensively studying how SCFAs are associated with
colorectal tumors.

To accomplish this task we directly measured the concentration of
acetate, butyrate, and propionate within fecal samples for two different
groups. The first group had a sample obtained at a single cross
sectional point in time while the second group had samples obtained
before (pre-) and after (post-)treatment for colorectal tumors. To
provide further support for our SCFA findings we also used PICRUSt (21)
and metagenomic sequencing to investigate differences in genes encoding
enzymes involved with SCFA synthesis between individuals without colon
tumors, patients with adenomas, and patients with carcinomas. Next, we
investigated whether taxa associated with SCFA production were important
to disease classification models. First, using the cross-sectional data,
we analyzed the number of correlations between OTU relative abundance
and SCFA concentrations across individuals without colon tumors and
patients with adenomas or carcinomas. Second, we assessed the affect
adding SCFA concentrations to OTU data had on classification of patients
with adenomas or carcinomas using the Random Forest algorithm (22).
Third, we analyzed how well 16S rRNA gene sequencing predicts SCFA
concentrations. Finally, we compared whether SCFA concentrations replace
important taxa in disease models and whether the same OTUs are the most
important variables used to classify disease and SCFA concentration.
Collectively, this investigation provides additional information as to
whether SCFAs are decreased in patients with colorectal tumors and
provides context as to whether targeting taxa to increase SCFA
concentrations is a viable option to protect against colon
tumorigenesis.

@@

\textbf{Gene abundance for enzymes involved in SCFA synthesis are the
same for individuals without colon tumors and patients with adenomas or
carcinomas.} In order to provide further confirmation and support of our
SCFA concentration results we investigated the genes encoding specific
enzymes involved with SCFA synthesis {[}Table S1{]}. Using this list of
specific genes {[}Table S1{]}, we looked for differences in gene
abundance between individuals without colon tumors and patients with
adenomas or carcinomas. Although we intended to analyze all the genes in
the list {[}Table S1{]}, not all of the KEGG genes were identified
during our analysis. We first analyzed imputed gene relative abundance,
calculated from our OTU data that was generated from 16S rRNA gene
sequencing. We found no difference in any of the imputed gene relative
abundances for enzymes involved with acetate, butyrate, or propionate
synthesis (P-value \textgreater{} 0.90) {[}Table S2{]}. Since butyrate
is one of the most studied SCFAs in the context of CRC, we visualized
the observed lack of difference between indviduals without colon tumors
and patients with adenomas or carcinomas in gene abundance for enzymes
involved with SCFA synthesis using the butyrate kinase gene {[}Figure
2A{]}. Additionally, using a paired Wilcoxon rank-sum test, there also
was no difference in imputed gene relative abundance between pre- and
post-treatment samples for any genes involved with SCFA synthesis
(P-value \textgreater{} 0.70) {[}Table S3{]}. Next, we took a subset of
these 490 fecal samples (n=85) and used metagenomic sequencing to
confirm these results. Like the imputed gene results, metagenomic
analysis found that there was no difference in any of the genes involved
in SCFA synthesis between individuals without colon tumors (n=29) and
patients with adenomas (n=28) or carcinomas (n=28) (P-value
\textgreater{} 0.70) {[}Table S4{]}. This similarity between indviduals
without colon tumors and patients with adenomas or carcinomas is
highlighted again by visualizing the gene abundance for the butyrate
kinase gene {[}Figure 2B{]}. These observations provide evidence that
gene prevalence for enzymes involved in SCFA synthesis does not change
due to colorectal tumors and provides further support for our original
SCFA concentration observations.

\textbf{The number of OTUs positively associated with SCFA concentration
were similar between individuals without colon tumors and patients with
adenomas or carcinomas.} Having found no difference between individuals
without colon tumors and patients with adenomas or carcinomas in SCFA
concentrations or genes encoding enzymes involved with SCFA synthesis,
we next investigated if specific OTUs correlated with SCFA
concentrations. The main goal of this analysis was to identify if there
were OTUs that were significantly associated with SCFA concentrations
and if this was different between individuals that did not have colon
tumors, had an adenoma, or had a carcinoma. To accomplish this we used
Spearmans rho, a non-parametric measure of association, and tested if
there was a correlation that was signficantly greater than zero. We
found that taxa from \emph{Clostridiales}, \emph{Lachnospiraceae}, and
\emph{Ruminococcaceae} dominated statistically significant OTU
correlations {[}Figure 3 \& Table S5{]}. There was a noticeably higher
number of significant negative correlations associated with patients
with adenomas for all SCFAs tested {[}Figure 3{]}. In particular, OTUs
from the \emph{Ruminococcaceae} family had the largest share of these
negative correlations within patients with adenomas {[}Figure 3{]}.
Patients with adenomas also had more positive correlations between OTUs
and SCFA concentrations, but their total number was more similar to
individuals without colon tumors or patients with carcinomas versus the
analogous comparison for the number of negative correlations {[}Figure
3{]}. Additionally, the number of positive correlations between OTUs and
SCFA concentrations was similar between individuals without colon tumors
and patients with a carcinoma {[}Figure 3{]}. Finally, when we used
high/low SCFA groups based on the overall median concentration for each
SCFA instead of SCFA concentrations a similar pattern was still observed
{[}Figure S1 \& Table S6{]}. Overall, these results suggest that the
resident taxa that may change the most due to colon tumors may not be
ones that are responsible for the production of acetate, butyrate, or
propionate.

\textbf{SCFA concentrations do not replace important
\emph{Clostridiales}, \emph{Lachnospiraceae}, and \emph{Ruminococcaceae}
OTUs in Random Forest models built to classify tumors.} Despite the lack
of difference in positive correlations between OTUs and SCFA
concentrations between individuals with and without tumors, OTUs
associated with SCFA concentrations could still be the most important
variables to Random Forest models built to classify patients with
adenomas or carcinomas. We tested this by using the Random Forest
algorithm to build models with OTU abundance data or OTU abundances and
SCFA concentrations to classify normal versus adenoma and normal versus
carcinoma fecal samples. With these models we compared whether OTUs with
taxonomic classification to \emph{Clostridiales},
\emph{Lachnospiraceae}, and \emph{Ruminococcaceae} remained when SCFA
concentrations were added to the model. Additionally, we also compared
whether any of the important OTUs that remained also had significant
correlations with SCFA concentrations. Both our adenoma and carcinoma
models classified patients with a similar degree of success, as measured
by the area under the curve (AUC) (P-value \textgreater{} 0.05)
{[}Figure 4A \& 4D{]}. After the addition of SCFA concentrations to the
adenoma or carcinoma models, many OTUs with taxonomic classification to
\emph{Clostridiales}, \emph{Lachnospiraceae}, and \emph{Ruminococcaceae}
remained as important variables to the model {[}Figure Figure 4B-C \&
4E-F{]}. After adding SCFA concentrations to the adenoma model, there
were only 2 OTUs significantly associated with SCFA concentration that
were part of the top 10 most important variables and both were postively
associated with acetate, butyrate, or propionate concentrations. When
SCFA concentrations were added to the carcinoma model, only 1 OTU that
was associated with SCFA concentration was in the top 10 most important
variables and it was negatively associated with acetate and butyrate
concentrations. In combination with the previous results on OTU
correlations, these observations provide additional evidence that the
resident taxa that are associated with protection against tumorigeneis
are not ones associated with acetate, butyrate, or propionate
production.

\textbf{The most important OTUs for SCFA concentration and tumor Random
Forest models are not the same.} It is possible that due to the way the
Random Forest algorithm works, OTUs associated with SCFA concentrations
could be downweighted in importance within the adenoma or carcinoma
models when SCFA concentrations are included. To test if this is the
case we used OTU data and built Random Forest models to classify SCFA
concentrations. Overall, the correlation between the predicted and
actual SCFA concentrations were moderately associated with each other
{[}Figure 5A{]}. Additionally, because the training set
R\textsuperscript{2} was always higher than the test set
R\textsuperscript{2}, all SCFA concentration models tended to be over
fit, suggesting that rarer taxa were important for these classifications
{[}Figure 5A{]}. There also was a difference in accuracy based on
whether the fecal sample was from an individual without colon tumors or
from patients with adenomas or carcinomas {[}Figure 5B{]}. When
comparing the adenoma or carcinoma model to the SCFA concentration
models there was minimal overlap between these model's most important
OTUs {[}Figure 4B-C, 4E-F and 5C-E{]}. The only OTU that was in the top
10 most important variables and had overlap between the models was
OTU00167 (\emph{Clostridiales}) {[}Figure 4B-C, 4E-F, 5C-E{]}. Similar
observations were made when using high/low SCFA groups based on the
median SCFA concentration {[}Figure S2{]}. Collectively, these
observations provide evidence that it is possible to identify specific
OTUs associated with higher SCFA concentrations and accordingly these
OTUs belong to taxa known to produce acetate, butyrate, and propionate.
Although it is possible to identify OTUs associated with SCFA
production, overall, our results do not support the hypothesis that SCFA
concentration or OTUs associated with their production are different
between individuals with no colon tumors and patients with adenomas or
carcinomas.

@@

The observations from this study do not support the hypothesis that SCFA
concentrations are different in individuals with tumors. Whether we
directly measured the SCFA concentration or investigated genes that
encoded enzymes used for their production, no difference could be
identified between individuals without tumors and patients with adenomas
or carcinomas {[}Figure 1 \& 2{]}. There were differences in the number
of significant correlations between SCFA concentration and OTU relative
abundance based on whether individuals did not have colon tumors, had an
adenoma, or had a carcinoma. However, SCFA concentrations did not
provide increased model accuracy for tumor classification {[}Figure 3-4
\& S1{]}. In models with SCFA concentrations included, many OTUs that
classified to \emph{Clostridiales}, \emph{Lachnospiraceae}, and
\emph{Ruminococcaceae} remained as important variables for the model
{[}Figure 4{]}. Additionally, when models using OTU relative abundance
to classify SCFA concentrations were assessed, the OTUs that classified
to \emph{Clostridiales}, \emph{Lachnospiraceae}, and
\emph{Ruminococcaceae} were not the same as the OTUs that classified to
these taxa in the tumor models {[}Figure 4-5{]}. Collectively, our
observations suggest that the resident taxa from \emph{Clostridiales},
\emph{Lachnospiraceae}, and \emph{Ruminococcaceae} that are different
between individuals without colon tumors and patients with adenomas or
carcinomas, are not the same as those involved with SCFA production.

Although SCFAs have been shown to be anti-tumorigenic, most of these
studies have been performed in model systems (16, 17). Many of the
\emph{in vivo} studies use proxies such as fiber supplementation rather
than SCFAs directly (14). Although it is well known that breakdown
products from gut bacteria results in SCFA production, fiber effects on
tumorigenesis may be through other mechanisms in these \emph{in vivo}
model systems. Additionally, the observations in humans on the benefit
of SCFAs in preventing tumorigenesis have been mixed. In previous
case/control studies lower SCFA concentrations were observed in patients
with carcinomas versus those without carcinomas (19). Yet, this is in
contrast to multiple randomized-controlled trials that have found no
difference in tumor recurrence between patients who do and do not get
fiber supplementation (20, 23). In contrast to the \emph{in vivo} model
findings, the observations made in these randomized-controlled trials
would suggest that SCFAs do not prevent or slow tumorigenesis. One
reason for these results is that SCFA concentrations and responses to
fiber vary quite a bit between healthy individuals (24). This
information taken together with our observations would suggest that
either SCFAs provide little to no benefit as an anti-tumorigenic
compound in colorectal cancer or that individuals who do not respond to
fiber supplementation would need to acquire these bacteria to achieve a
benefit.

Another possible alternative explanation as to why no difference in SCFA
concentration between individuals without colon tumors and patients with
adenomas or carcinomas was observed, could be because only certain types
of colorectal cancers are affected by SCFAs. One limitation of current
research into the effect of SCFAs and the microbiota in CRC has been
that all tumors are treated as the same type. However, there are known
differences in the types of mutations that occur (25) and treating all
tumors as equal may actually hide any benefit that could be found in
certain subsets of individuals. Similar to the idea of using
immunotherapy as a targeted treatment option for specific tumors (26),
SCFAs may have beneficial effects for distinct types of colorectal
tumors. Future research will need to test if this is a valid hypothesis.
Regardless of this limitation, our results in combination to previous
randomized controlled trials on fiber supplementation (20, 23) suggests
that using SCFAs as a general treatment for colorectal cancer is
unlikely to provide a reduction in tumorigenesis.

@@

The potential protection against colorectal cancers may not be from
SCFAs, even though taxa associated with their production are higher in
individuals without tumors versus patients with carcinomas (10, 11, 27).
Our data is consistent with the literature despite finding no difference
in SCFA concentrations between individuals without colon tumors and
patients with adenomas or carcinomas. In particular, we observed that
the important OTUs in our tumor models were to taxa that are associated
with SCFA production but that these OTUs were not themselves associated
with acetate, butyrate, or propionate production. One of the clearest
results to highlight this within our study were the results showing that
different OTUs from the same taxonomic classification are in tumor and
SCFA Random Forest models {[}Figure 4 \& 5{]}. This leads to the
possibility that protection may be through other routes. For example,
there could be a different pathway or other less extensively studied
metabolites that provide the necessary protection against tumorigenesis.
Alternatively, previous research has shown that non-resident oral
microbes, such as \emph{Fusobacterium}, \emph{Porphyromonas},
\emph{Parvimonas}, and \emph{Peptostreptococcus} are associated with
carcinomas (6, 12, 13) and can increase inflammation (28). When these
non-resident oral microbes increase inflammation they change the
existing environment to one in which is more hospitable for themselves
(28). It is possible that these resident microbes provide protection not
through a metabolite but instead through active exclusion of these
non-resident oral microbes which in turn prevents them from establishing
themselves within the community. Overall, even though we did not find
lower concentrations of SCFAs associated with colorectal tumors, we
think that our results open up many new and exciting avenues for
investigation into how metabolites and the microbiota can prevent
tumorigenesis.
